3D Regression Heat Map Analysis of Population Study Data

### 2015-06-23 draft.tex
further cuts to fit the length specifications

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Interactive Visual Analysis, regression, Epidemiology, Breast Cancer, Hepatic Steatosis}

@@

Testing features for associations with diseases using regression models is one of the most important epidemiological tools.
%%
Using regression analysis to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
%%
Due to the amount of data and only limited overview visualizations, possible correlations may be missed.
%%
Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target feature.
%%
They mostly highlight correlations between features, which are known to the domain expert (e.g., correlation between body size and spine shape).
%%
We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis w.r.t. a specific disease or hypothesis.
%%
This is achieved by providing template regression formulas, which are applied to all potential feature combinations.
%%
Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
%%
Difference views between regression formulas allow to assess the influences of individual features on the process.

Our contributions are:

@@

Medical image data are also acquired and analyzed in modern population studies.
%%
We incorporate image-derived features, but do not focus on analyzing image data.
%%
More discussions of population study data types and characteristics can be found in the works of Preim et al.~\cite{Preim2014} and Toennies et al.~\cite{Toennies2015}.
%%
\\\\
Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
%%
The analysis model has to be adjusted by normalizing all included features w.r.t. the confounder.
%%
\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely for older subjects.
%%
It also influences the general body condition and thereby almost all features acquired through population studies.
%%
Another important confounder is \emph{gender}.
%%
Other confounders have to be selected by epidemiologists specific to the investigated condition.

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data and impose the basis of this work.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
\add{
%%
The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model and is suited for comparing models.
%%
A small AIC value indicates a higher quality model.
%%
}
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value ranges between $[0, 1]$, where 1 encodes a perfect fit.

\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
%%
\begin{equation}
Dependent \sim Independent_1 + ... + Independent_n
\label{eq:RegressionNotation}
\end{equation}
%%
An example of a regression formula would be $KidneyDisorder \sim Smoking + Obesity$.
The most commonly used regression operators comprise:
%%
\begin{itemize}
	\item $+,-$ inclusion/exclusion of the variable (e.g. $x\pm y$),
	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
	\item $*$ inclusion of the variables as well as their interactions (e.g. $x*y$)
	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
\end{itemize}
%%
The class of the target feature restricts the regression type.
%%
Different regression types are available, we focus on the following:
%%
\paragraph{Linear Regression for Continuous Target.}
%%
The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
%%
The \emph{dependent} variable has to be of \emph{continuous type}.
%%
\add{
%%
Linear regression models can also be described using the \emph{adjusted} $R^2$, which considers the number of dependent features.
%%
It yields lower values, when features with little entropy w.r.t. the target feature are included.
%Additional metrics describing linear regression models comprise of the \emph{adjusted} $R^2$, which also adjusts for the number of dependent features and yields lower values, when features with little entropy towards the target feature are included.
%%
The $f$-$statistic$ determines how much improvement the independent variables will add to the model compared to them being not considered at all.
%%
}

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
%%
Comparisons have to be handled with care.
%%
\subsection{The Study of Health in Pomerania (SHIP)}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
%It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data per participant.
%%
A second cohort, \texttt{SHIP-TREND} was started in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.
%%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
Tukey already stated in 1977 that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to \emph{derive} hypotheses, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
%The work of M\"{u}hlbacher et al.~\cite{Muehlbacher}, Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
\paragraph{Parameter space analysis using regression models.}
%%
\magenta{
Sedlmair et al.~\cite{Sedlmair} present a conceptual taxonomy of parameter space analysis.
%%
%It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
%%
Using the taxonomy, approaches can be classified to find works with similar problems and derive new design ideas and evaluate ideas.
%The taxonomy allows to classify approaches to find works with similar problems to derive new design ideas and evaluate ideas.
%%
Based on their input parameter taxonomy, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
%%
We solve a \emph{fitting} task by aiming to find models well suited for describing the input data.
}

\magenta{
%% A Partition-based Framework for Building and Validating Regression Models
The approach of M\"{u}hlbacher et al.~\cite{Muehlbacher} is closest to ours.
%%
They provide a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
%%
%Correlations between features with the target yield a ordered matrix-plot, where feature combinations are used to depict models of interest.
%A view for single-relationships with the target and single features as well as a matrix-plot for feature combinations are used to depict models of interest.
%%
%The visualization allows to assess different model complexities.
%%
Existing regression models can be validated and compared using 3D views and 2D slice-views.
%%
We employ the same steps by preprocessing features using correlations with the target and providing an overview over combinations.
%%
M\"{u}hlebacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot-matrix view, while we cover more features by abstracting the models.
%%
%We also take categorical targets and regression formula inputs into account.
%%
}

@@

\paragraph{Visual analysis of epidemiological data.}
%%
Zhang et al.~\cite{Zhang2014, Gotz2014} is closest to our work regarding the application of visual analysis on cohort study data.
%%
They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} (and its manipulation using operations), \emph{views} and \emph{analytics}.
%%
They use the system to find longitudinal pathways for diseases on the basis of health records.
%%
Sub-groups are automatically created by incorporating \emph{analytics} and through selection in expert-guided \emph{views}.
%%
%They do not go into the detail w.r.t the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
%%
%Their data differs strongly from ours, having multiple moments for each subject, yielding a approach focused on finding pathways of interest.
%%
We use their requirements as guidelines, which incorporate \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
%%
%We implement this by supporting both hypothesis-based and hypothesis-free analysis.
%%
Angelelli et al.~\cite{Angelelli} visualize image-derived an non-image data using cube data structures with focus on comparison and knowledge extraction.
%%
Their proposed engine represents multiple heterogenous cohort study data sets as normalized data cubes.
%%
%Thus, data redundancy is avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
%%
%This allows for saving memory space.
%%
%Their presented aggregation functions allow for a fast data calculation to map it using different linked visualizations, such as scatter plots, bar charts or histograms.
%%
They use \emph{Pearsson's} $r$ to characterize relationships with target features and employ list views and scatter plots to visualize and rank them.
%%
%We use regression models and associated $R^2$ values to depict relationships between a target feature and multiple independent features instead of one and employ the information into one global visualization.
%%
In contrast, we aim for a fast large scale correlation analysis incorporating many features to derive insight into the data.
%%
%\paragraph{Visual Analysis in Public Health.}
%%--
% Shneiderman et al.~\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data. %, e.g. using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
% %%
% %They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
% %%
% %We focus on the latter.
% %%
% %Shneiderman and colleagues also present challenges,
% The presented challenges in this application require the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
% %%
% Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
%%
%We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
%%

@@

%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
Niemann et al.~\cite{Niemann2014} investigate risk factors of hepatic steatosis using decision trees with interactive data mining tools.
%%
They extract classification rules that serve as basis for our proof-of-concept tests.
%%
Niemann et al.~\cite{Niemann2015} improved the classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
%%
We try to reproduce their results and investigate findings presented by Niemann et al.~\cite{Niemann2014} further.
%%
\paragraph{Prior work.}
In our prior work, we analyzed the healthy aging process of the lumbar spine.
%%
%We utilized semi-automatic algorithms do detect the lumbar spine shape \cite{Klemm2013VMV, Rak2013}.
%%
%The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
%%
We defined an Interactive Visual Analysis workflow for image-centric population study data \cite{Klemm2014VIS}.
%%
We extended the feature selection step (recall Sec. \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods.
%%
Information visualizations of population study features were augmented with extracted image data, yielding linked views with both image and non-image information.
%%
The 2D heat map \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
%%
Its great popularity among our domain experts was the inspiration for this work.
%%

While the presented techniques were well received by the epidemiological experts, the explanatory power w.r.t. back pain was limited.

@@

%\subsection{Abstracting Regression Results Using $R^2$}
\subsection{Abstracting Regression Results}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/heat_map}
 \caption{
 (a) Overview visualization using a 2D heat map of the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The descriptive metrics extracted from the regression model are mapped to color saturation (for $R^2$, saturated color shows a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heat map visualizations.
 %%
 These represent the slices in our \emph{3D Regression Heat Map}.
 %%
 The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Heat_Map}
\end{figure}
%%
\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a 2D heat map.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{3D Regression Heat Map} analysis we are interested in the quality-of-fit of the resulting model.
%%
This allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment.
%%
% \add{
% %%
% Users with in-depth knowledge about regression models are given the option to render the \emph{3D Regression Heat Map} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
% %%
% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
% %%
% To avoid shutting out users unfamiliar with these metrics, the \emph{3D Regression Heat Map} is set per default to show $R^2$ values.
% }
%%
\paragraph{2D (slice) view.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same 2D heat map by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Heat_Map}a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%
\design{
%%
Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
%%
Therefore, these experts can investigate the heat map with emphasis on specific aspects of the model.
%%
}
%%
\add{
%%
$Adjusted$ $R^2$ can be represented in the same way, since they share the scale.
%%
$AIC$ values are not restricted to a normalized scale.
%%
They have to be normalized in order to map it onto color saturation.
%%
%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
The resulting scale may be distorted by outliers derived from poor regression models.
%%
To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
%%
Outliers can be cut off to emphasize ranges of interest.
%%
Small $AIC$ values indicate a good model.
%%
Hence, we inverted the transfer function mapping it to color, mapping low $AIC$ features to saturated colors.
%%
}
\design{
%%
To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
%%
}

\paragraph{3D view.}
%%
Introducing $Z$ creates a 3D heat map (Fig.~\ref{fig:Heat_Map}b).
%%
\add{The selected metric (per default set to $R^2$)} of each heat map entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
%%
\magenta{Object size is not used to encode information because it would result in a cluttered view.}
%%
\design{
%%
Epidemiologists with background in statistics argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
Therefore, we decided to map metrics of different regression methods on distinct colors (i.e., orange for linear regression and blue for logistic regression).
%%
}
%%
%The visualization of descriptive metrics derived from different 3D Regression Heat Maps (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
%Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., orange for linear regression and blue for logistic regression).
%%
Thus, the visualization can be easily extended using other regression types.
%%
For \emph{3D Regression Heat Maps} with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
%%
\magenta{
As mentioned previously, the dimension reduction using the CFS algorithm potentially removes features, which are of interest for the domain expert.
%%
To tackle this problem, the $z$ dimension of the visualization contains all features of the data set, allowing to assess their influence on the models.
%%
The $x$ and $y$ dimensions are restricted to the features extracted from the CFS algorithm.
}
%%
\\\\
%%
Our goal is to create an overview visualization for a data set.
%%
We also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
The different analysis approaches require different starting points using the \emph{3D Regression Heat Map}.

>>>>

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Interactive Visual Analysis, Regression Analysis, Heat Map, Epidemiology, Breast Cancer, Hepatic Steatosis}

@@

Testing features for associations with diseases using regression models is one of the most important epidemiological tools.
%%
Using regression analysis to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
%%
Due to the amount of data and only limited overview visualizations, possible correlations may be missed.
%%
Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target feature.
%%
They mostly highlight correlations between features, which are known to the domain expert (e.g., correlation between body size and spine shape).
%%
We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis w.r.t. a specific disease or hypothesis.
%%
Therefore we provide template regression formulas, which are applied to all potential feature combinations.
%%
Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
%%
Difference views between regression formulas allow to assess the influences of individual features on the process.
%%
Our contributions are:

@@

Medical image data are also acquired and analyzed in modern population studies.
%%
We incorporate image-derived features, but do not focus on analyzing image data.
%%
More discussions of population study data types and characteristics can be found in the works of Preim et al.~\cite{Preim2014} and Toennies et al.~\cite{Toennies2015}.
%%

Features influencing the exposure as well as the outcome of an analysis are called \emph{confounders} and have to be specially treated.
%%
The analysis model has to be adjusted by normalizing all included features w.r.t. the confounder.
%%
\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely for older subjects.
%%
It also influences the general body condition and thereby almost all features acquired through population studies.
%%
Another important confounder is \emph{gender}.
%%
Confounders have to be selected by epidemiologists specific to the investigated condition.
%%
\paragraph{The Study of Health in Pomerania (SHIP).}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
%It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data per participant.
%%
A second cohort, \texttt{SHIP-TREND} was started in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
%For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
Most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data and impose the basis of this work.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
\add{
%%
The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model and is suited for comparing models.
%%
A small AIC value indicates a higher quality model.
%%
}
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value ranges between $[0, 1]$, where 1 encodes a perfect fit.
%%

\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
%%
$Dependent \sim Independent_1 + ... + Independent_n$.
% \begin{equation}
% Dependent \sim Independent_1 + ... + Independent_n
% \label{eq:RegressionNotation}
% \end{equation}
%%
An example of a regression formula would be $KidneyDisorder~\sim~Smoking~+~Obesity$.
The most commonly used regression operators comprise:
%%
\begin{itemize}
	\item $+,-$ inclusion/exclusion of the variable (e.g. $x\pm y$),
	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
	\item $*$ inclusion of the variables as well as their interactions (e.g. $x*y$)
	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
\end{itemize}
%%
The class of the target feature restricts the regression type.
%%
Different regression types are available, we focus on the following:
%%
\paragraph{Linear Regression for Continuous Target.}
%%
The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
%%
The \emph{dependent} variable has to be of \emph{continuous type}.
%%
\add{
%%
Linear regression models can also be described using the \emph{adjusted} $R^2$, which considers the number of dependent features.
%%
It yields lower values, when features with little entropy w.r.t. the target feature are included.
%Additional metrics describing linear regression models comprise of the \emph{adjusted} $R^2$, which also adjusts for the number of dependent features and yields lower values, when features with little entropy towards the target feature are included.
%%
The $f$-$statistic$ determines how much improvement the independent variables will add to the model compared to them being not considered at all.
%%
}

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
%%
Comparisons have to be handled with care.
%%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
Tukey already stated in 1977 that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to \emph{derive} hypotheses, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
%The work of M\"{u}hlbacher et al.~\cite{Muehlbacher}, Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
\paragraph{Parameter space analysis using regression models.}
%%
\magenta{
Sedlmair et al.~\cite{Sedlmair} present a conceptual taxonomy of parameter space analysis.
%%
%It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
%%
%Using the taxonomy, approaches can be classified to find works with similar problems and derive new design ideas and evaluate ideas.
%The taxonomy allows to classify approaches to find works with similar problems to derive new design ideas and evaluate ideas.
%%
Based on their input parameter taxonomy, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
%%
We solve a \emph{fitting} task by aiming to find models well suited for describing the input data.
}

\magenta{
%% A Partition-based Framework for Building and Validating Regression Models
The approach of M\"{u}hlbacher et al.~\cite{Muehlbacher} is closest to ours.
%%
They provide a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
%%
%Correlations between features with the target yield a ordered matrix-plot, where feature combinations are used to depict models of interest.
%A view for single-relationships with the target and single features as well as a matrix-plot for feature combinations are used to depict models of interest.
%%
%The visualization allows to assess different model complexities.
%%
Existing regression models can be validated and compared using 3D views and 2D slice-views.
%%
%We employ the same steps by preprocessing features using correlations with the target and providing an overview over combinations.
%%
M\"{u}hlbacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot-matrix view, while we cover more features by abstracting the models.
%%
%We also take categorical targets and regression formula inputs into account.
%%
}

@@

\paragraph{Visual analysis of epidemiological data.}
%%
Zhang et al.~\cite{Zhang2014, Gotz2014} is closest to our work regarding the application of visual analysis on cohort study data.
%%
They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} (and its manipulation using operations), \emph{views} and \emph{analytics}.
%%
They use the system to find longitudinal pathways for diseases on the basis of health records.
%%
%Sub-groups are automatically created by incorporating \emph{analytics} and through selection in expert-guided \emph{views}.
%%
%They do not go into the detail w.r.t the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
%%
%Their data differs strongly from ours, having multiple moments for each subject, yielding a approach focused on finding pathways of interest.
%%
We apply their requirements as guidelines, which incorporate \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
%%
%We implement this by supporting both hypothesis-based and hypothesis-free analysis.
%%
Angelelli et al.~\cite{Angelelli} visualize image-derived an non-image data using cube data structures with focus on comparison and knowledge extraction.
%%
%Their proposed engine represents multiple heterogenous cohort study data sets as normalized data cubes.
%%
%Thus, data redundancy is avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
%%
%This allows for saving memory space.
%%
%Their presented aggregation functions allow for a fast data calculation to map it using different linked visualizations, such as scatter plots, bar charts or histograms.
%%
They use \emph{Pearsson's} $r$ to characterize relationships with target features and employ list views and scatter plots to visualize and rank them.
%%
%We use regression models and associated $R^2$ values to depict relationships between a target feature and multiple independent features instead of one and employ the information into one global visualization.
%%
In contrast, we aim for a fast large scale correlation analysis incorporating many features to derive insights.
%%
%\paragraph{Visual Analysis in Public Health.}
%%--
% Shneiderman et al.~\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data. %, e.g. using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
% %%
% %They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
% %%
% %We focus on the latter.
% %%
% %Shneiderman and colleagues also present challenges,
% The presented challenges in this application require the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
% %%
% Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
%%
%We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
%%

@@

%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
Niemann et al.~\cite{Niemann2014} investigate risk factors of hepatic steatosis using decision trees with interactive data mining tools.
They extract classification rules that serve as basis for our proof-of-concept tests.
Niemann et al.~\cite{Niemann2015} improved the classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
We try to reproduce their results and investigate findings presented by Niemann et al.~\cite{Niemann2014} further.
\paragraph{Prior work.}
In our prior work, we analyzed the healthy aging process of the lumbar spine.
%We utilized semi-automatic algorithms do detect the lumbar spine shape \cite{Klemm2013VMV, Rak2013}.
%The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
We defined an Interactive Visual Analysis workflow for image-centric population study data \cite{Klemm2014VIS}.
We extended the feature selection step (recall Sec. \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods.
Information visualizations of population study features were augmented with extracted image data, yielding linked views with both image and non-image information.
The 2D heat map \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
Its great popularity among our domain experts was the inspiration for this work.
The presented techniques were well received by the epidemiological experts, but the explanatory power w.r.t. back pain was limited.

@@

%\subsection{Abstracting Regression Results Using $R^2$}
\subsection{Abstracting Regression Results}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/heat_map}
 \caption{
 (a) Overview visualization using a 2D heat map of the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The $R^2$ metrics extracted from the regression model are mapped to color saturation (a saturated color indicates a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heat maps.
 %%
 These represent the slices in our \emph{3D Regression Heat Map}.
 %%
 The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Heat_Map}
\end{figure}
%%
\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a 2D heat map.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{3D Regression Heat Map} analysis we are interested in the quality-of-fit of the resulting model.
%%
This allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment.
%%
% \add{
% %%
% Users with in-depth knowledge about regression models are given the option to render the \emph{3D Regression Heat Map} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
% %%
% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
% %%
% To avoid shutting out users unfamiliar with these metrics, the \emph{3D Regression Heat Map} is set per default to show $R^2$ values.
% }
%%
\paragraph{2D (slice) view.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same 2D heat map by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Heat_Map}a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%
\design{
%%
Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
%%
Therefore, these experts can investigate the heat map with emphasis on specific aspects of the model.
%%
}
%%
\add{
%%
$Adjusted$ $R^2$ can be represented in the same way, since they share the scale.
%%
$AIC$ values are not restricted to a normalized scale.
%%
They have to be normalized in order to map it onto color saturation.
%%
%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
The resulting scale may be distorted by outliers derived from poor regression models.
%%
To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
%%
Outliers can be cut off to emphasize ranges of interest.
%%
Small $AIC$ values indicate a good model.
%%
Hence, we inverted the transfer function mapping it to color, mapping low $AIC$ features to saturated colors.
%%
}
\design{
%%
To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
%%
}

\paragraph{3D view.}
%%
Introducing $Z$ creates a 3D heat map (Fig.~\ref{fig:Heat_Map}b).
%%
\add{The selected metric (per default set to $R^2$)} of each heat map entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
%%
\magenta{Object size is not used to encode information because it would result in a cluttered view.}
%%
\design{
%%
Epidemiologists with background in statistics argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
Therefore, we decided to map metrics of different regression methods on distinct colors (i.e., orange for linear regression and blue for logistic regression).
%%
}
%%
%The visualization of descriptive metrics derived from different 3D Regression Heat Maps (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
%Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., orange for linear regression and blue for logistic regression).
%%
Thus, the visualization can be easily extended using other regression types.
%%
For \emph{3D Regression Heat Maps} with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
%%
\magenta{
As mentioned previously, the dimension reduction using the CFS algorithm potentially removes features important to the domain expert.
%%
To tackle this problem, the $z$ dimension of the visualization contains all features of the data set, allowing to assess their influence.
%%
The $x$ and $y$ dimensions are restricted to the features extracted from the CFS algorithm.
}
%%
\\\\
%%
Our goal is to create an overview visualization for a data set.
%%
We also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
The different analysis approaches require different starting points using the \emph{3D Regression Heat Map}.
